Distinguishing Spin-Aligned and Isotropic Black Hole Populations With Gravitational Waves

### 2017-04-12 paper/AlignedIsotropicGW.tex
Citations, oh so many citations, in the abstract.

\begin{abstract}
  The first direct detections of \acp{GW} from merging \acp{BBH} open
  a unique window into the \ac{BBH} formation environment.  One
  promising signature of the formation environment is the angular
  distribution of the \ac{BH} spins; systems formed through dynamical
  interactions among already-compact objects are expected to have
  isotropic spin orientations whereas binaries formed from pairs of
  stars born together are more likely to have spins preferentially
  aligned with the binary orbit as a consequence of their joint
  evolution toward a \ac{BBH} system.  We consider existing \ac{GW}
  measurements of $\chieff$, the best-measured combination of spin
  parameters, in the three likely binary black hole detections
  GW150914, LVT151012, and GW151226.  If binary black hole spin
  magnitudes extend to high values, as is suggested by observations of
  black hole X-ray binaries, we show that the data already exhibit a
  $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
  ratio\footnote{An odds ratio of $r$ with $r \ll 1$ is equivalent to
    $x \sigma$ with $x = \Phi^{-1}\left( 1 - r/2 \right)$, where
    $\Phi$ is the unit normal CDF.})  preference for an isotropic
  angular distribution.  By considering the effect of an additional 10
  detections drawn from the various models in the suite we show that
  if all observations come from a single population such an augmented
  data set would enable at least a $\OTwoSigmaIsoAlignedMin\sigma$
  ($\OTwoOddsIsoAlignedMin$ odds ratio) distinction between the
  isotropic and aligned models for the assumed spin magnitude
  distributions, and in most cases better than $5\sigma$
  ($2.9 \times 10^{-7}$ odds ratio).  The existing preference for
  either an isotropic spin distribution or low spin magnitudes for the
  observed systems will be confirmed (or overturned) confidently in
  the near future by subsequent observations.
\end{abstract}

>>>>

\begin{abstract}
  The first direct detections of \acp{GW}
  \citep{2016PhRvL.116f1102A,2016PhRvL.116x1103A,O1-BBH} from merging
  \acp{BBH} open a unique window into the \ac{BBH} formation
  environment.  One promising signature of the formation environment
  is the angular distribution of the \ac{BH} spins; systems formed
  through dynamical interactions among already-compact objects are
  expected to have isotropic spin orientations
  \citep{SigurdssonHernquist:1993,1993Natur.364..421K,PZMcMillan:2000,Rodriguez:2015,Stone:2016,2016ApJ...832L...2R}
  whereas binaries formed from pairs of stars born together are more
  likely to have spins preferentially aligned with the binary orbit as
  a consequence of their joint evolution toward a \ac{BBH} system
  \citep{TutukovYungelson:1973,TutukovYungelson:1993,Lipunov:1997,2016Natur.534..512B,Stevenson:2017,MandeldeMink:2016,Marchant:2016}.
  We consider existing \ac{GW} measurements of $\chieff$, the
  best-measured combination of spin parameters
  \citep{2016PhRvL.116x1102A,O1-BBH}, in the three likely binary black
  hole detections GW150914, LVT151012, and GW151226.  If binary black
  hole spin magnitudes extend to high values, as is suggested by
  observations of black hole X-ray binaries
  \citep{2015PhR...548....1M}, we show that the data already exhibit a
  $\OOneSigmaIsoAligned\sigma$ ($\OOneOddsIsoAligned$ odds
  ratio\footnote{An odds ratio of $r$ with $r \ll 1$ is equivalent to
    $x \sigma$ with $x = \Phi^{-1}\left( 1 - r/2 \right)$, where
    $\Phi$ is the unit normal CDF.})  preference for an isotropic
  angular distribution.  By considering the effect of an additional 10
  detections \citep{2016ApJ...833L...1A} drawn from the various models
  in the suite we show that if all observations come from a single
  population such an augmented data set would enable at least a
  $\OTwoSigmaIsoAlignedMin\sigma$ ($\OTwoOddsIsoAlignedMin$ odds
  ratio) distinction between the isotropic and aligned models for the
  assumed spin magnitude distributions, and in most cases better than
  $5\sigma$ ($2.9 \times 10^{-7}$ odds ratio).  The existing
  preference for either an isotropic spin distribution or low spin
  magnitudes for the observed systems will be confirmed (or
  overturned) confidently in the near future by subsequent
  observations.
\end{abstract}
